### 2024-07-29 cluster-analysis.qmd
20240728 - fixes
20240729 - overview of cluster analysis

```{r}

```

@@

```{r}
recentSeason <- max(player_stats_seasonal_offense$season, na.rm = TRUE)
recentSeason

@@

#### Fit the Cluster Model to the Optimal Number of Tiers {#sec-clusterAnalysisExampleNumTiers}

@@

@@

@@

>>>>

Whereas [factor analysis](#sec-factorAnalysis) evaluates how *variables* do or do not hang together—in terms of their associations and non-associations, cluster analysis evaluates how *people* are or or not similar—in terms of their scores on one or more variables.
The goal of cluster analysis is to identify distinguishable subgroups of people.
The people within a subgroup are expected to be more similar to each other than they are to people in other subgroups.
For instance, we might expect that there are distinguishable subtypes of Wide Receivers: possession, deep threats, and slot-type Wide Receivers.
Possession Wide Receivers tend to be taller and heavier, with good hands who catch the ball at a high rate.
Deep threat Wide Receivers tend to be fast.
Slot-type Wide Receivers tend to be small, quick, and agile.
In order to identify these clusters of Wide Receivers, we might conduct a cluster analysis with variables relating to the players' height, weight, percent of (catchable) targets caught, air yards received, and various metrics from the National Football League (NFL) Combine, including their times in the 40-yard dash, 20-yard shuttle run, and three cone drill.

There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, etc.
In this chapter, we focus on examples using model-based clustering with the `mclust` package, which uses Gaussian finite mixture modeling.

@@

```{r}
recentSeason <- max(player_stats_seasonal_offense$season, na.rm = TRUE) # also works: nflreadr::most_recent_season()
recentSeason

@@

#### Fit the Cluster Model to the Optimal Number of Tiers {#sec-clusterAnalysisExampleModel}

##### Quarterbacks {#sec-clusterAnalysisExampleModelQBs}

@@

##### Running Backs {#sec-clusterAnalysisExampleModelRBs}

@@

##### Wide Receivers {#sec-clusterAnalysisExampleModelWRs}

@@

##### Tight Ends {#sec-clusterAnalysisExampleModelTEs}

### 2024-07-29 player-evaluation.qmd
20240729 - overview of cluster analysis

Many aspects of a player's athletic profile, including tests of speed (40-yard dash), strength (bench press), agility (2-yard shuttle run; three cone drill), and jumping ability (vertical jump; broad jump) are available from the National Football League (NFL) Combine, which is especially relevant for evaluating rookies.
We demonstrate how to import data from the NFL Combine in @sec-combine.
There are also calculators that integrate information about body shape and information from the NFL Combine to determine a player's relative athletic score (RAS) for their position: <https://ras.football/ras-calculator/>

>>>>

Many aspects of a player's athletic profile, including tests of speed (40-yard dash), strength (bench press), agility (20-yard shuttle run; three cone drill), and jumping ability (vertical jump; broad jump) are available from the National Football League (NFL) Combine, which is especially relevant for evaluating rookies.
We demonstrate how to import data from the NFL Combine in @sec-combine.
There are also calculators that integrate information about body shape and information from the NFL Combine to determine a player's relative athletic score (RAS) for their position: <https://ras.football/ras-calculator/>
